---
output: html_notebook
---

<br>
<b><center><h1> Statistics Project II </h1></center></b>
<center><h3>by Sam Sharp, Alvin Ugochukwu, and Jon Scheaffer</h3></center>
<br>
<p>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;World Happiness Score is obtained by statistically calculating many different variables which describe aspects of society in a given country. These variables are GDP, social support, health/life expectancy, freedom, generosity, and perceptions of corruption. The dataset used is from Kaggle (https://www.kaggle.com/datasets/ajaypalsinghlo/world-happiness-report-2022) and uses data from the World Gallup Poll. Importantly, none of the predictor variables appear to be collinear, which strengthens the argument for the use of multiple linear regression. The dataset consists of entries for 146 countries, using 6 predictors that each linearly correlate with the World Happiness Score.</p>
<br>


```{r, echo=FALSE, results='hide'}
library(MASS) 
library(reshape2) 
library(reshape)
library(plotly)
library(car)
```

```{r, echo=FALSE}
data <- read.csv("Resources/World Happiness Report 2022.csv")
data

hs = data$Happiness.score
gdp = data$Explained.by..GDP.per.capita
soc = data$Explained.by..Social.support
hle = data$Explained.by..Healthy.life.expectancy
free = data$Explained.by..Freedom.to.make.life.choices
gene = data$Explained.by..Generosity
poc = data$Explained.by..Perceptions.of.corruption

Wh = data$Whisker.high
wl = data$Whisker.low
dys = data$Dystopia..1.83....residual
rank = data$RANK
```
<p>Above:The table displaying the data in the World Happiness report.</p>
------------------------------------------------------
<br>

```{r, echo=FALSE}
df <- data.frame(hs, soc, gdp, hle, free, gene, poc)
Variable <- c("hs", "soc", "gdp", "hle", "free", "gene", "poc")
Description <- c("Happiness score, the predicted result from independent variables", "The level of social programs and support available", "The GDP of the country in question", "The healthy life expectancy of a person in the country", "The ability of a person in the country to do what they wish", "The level of generosity someone experiences in the country", "The level of perceived corruption within institutions of power")
table <- data.frame(Variable, Description)
table
```
<p>Above:A table displaying the varaibles and there meanings.</p>
------------------------------------------------------
<br>

```{r, echo=FALSE, results='hide'}
summary(df)
```

```{r, echo=FALSE, results='hide'}
str(df)
```
<center><h2> Scatter Plots between the Different Variables </h2></center>
<br>
<div align="center">
```{r, echo=FALSE, fig.align = 'center'}
par(mfrow = c(1,3))
plot(hs, soc, pch = 19, col = "black")
plot(hs, gdp, pch = 19, col = "black")
plot(hs, hle, pch = 19, col = "black")
plot(hs, free, pch = 19, col = "black")
plot(hs, gene, pch = 19, col = "black")
plot(hs, poc, pch = 19, col = "black")
```
</div>

<center><h2> Creating the Linear Model </h2></center>
<br>

```{r, echo=FALSE}
model <- lm(hs ~ gdp + soc + hle + free + gene + poc)
summary(model)
```
<br>
<p>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;From the Estimate Std column we can derive the formula:</p>
```{r}
model2 <- lm(hs ~ gdp + soc + hle + free)
summary(model2)
```

<center><p style="font-size:25px"><b>hs = 1.67 + (0.55)X<sub>1</sub> + (1.41)X<sub>2</sub> + (1.27)X<sub>3</sub> + (1.60)<sub>4</sub> + (0.97)<sub>5</sub> + (0.73)X<sub>6</sub></b></p></center>
<br>

<p>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;The R squared value for this model is .7737 which tell us that 77% of the variation in the data is explained by the model. The overall F-statistic for the data is 79.2. The p-value is 2.2e-16, which is strong evidence to reject the null hypothesis that all the coefficients for each variable equal zero. All of this shows that the regression model is significant.</p>

------------------------------------------------------
<br>
<center><h2> Checking for Multicollinearity in the Data</h2></center>
```{r, echo=FALSE, results='hide'}
cormat <- round(cor(df),2)
melted_format <- melt(cormat)
```
<div align="center">
```{r, echo=FALSE}
ggplot(data = melted_format, aes(x=X1, y=X2, fill=value )) + 
  geom_tile() + geom_text(aes(label = value), color = 'white')
ggsave('Correlation_heatmap.png', width = 5, height = 5)
```
<p>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;This correlation heat map reveals that Generosity and Perception of Corruption are deeply uncorrelated to any of the other variables. </p>
</div>
```{r, echo=FALSE, results='hide'}
vif(model)
```
<div align="center">
```{r, echo=FALSE}
vif_values <- vif(model)
barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "steelblue")
```
```{r}
vif(model2)
```
```{r, echo=FALSE}
vif_values <- vif(model2)
barplot(vif_values, main = "VIF Values", horiz = TRUE, col = "steelblue")
```
```

<p>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;After running a multicollinearity check on the different independent variables, it is important to note that they do not appear to be strongly correlated at all as none of the variance inflation factors are even greater than 4.</p>
</div>
<center><h2> Conclusion </h2></center>
<br>
<p>&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;</p>

